### 2017-08-15 architecture.tex
added ref to foster paper and fixed various minor typos
data are plural

Selecting the policies, protocols, infrastructure, and
software settings to ensure that individual datasets are adequately secured requires consideration
on a case-by-case basis.
%To represent these policies, such that dataset owners can manage these considerations,
%we define safe-collections.
Safe collections enable system administrators to partition the key administrative
aspects of managing the relationship between data and users (e.g., analysts who use that data).
Safe collections encapsulate the data-store in which a dataset resides, and a consistent set of policies
that controls access to that dataset.

@@

\textbf{SecurityClass:} A binary classification that specifies the constraints
on accessing data in that collection. Setting the SecurityClass to \texttt{Public} means that the data is freely
(but securely) accessible to a set of permitted users and no other
data administration policies can be set.  Setting SecurityClass to \texttt{Private}
requires that safe collection policies are then set.

\textbf{Classification:} Categorizes the type of data stored in the
safe collection, including specifying that the data is controlled by
a license or that the data is subject to regulations (e.g., HIPAA or FEDRAMP).
This information is referenced and used by stewards and administrators when
managing a safe collection.

@@

Stewards are responsible for activities such as approving access to a collection, reviewing
export requests, and managing policies on the collection.
As described above, stewards are pre-associated with a particular safe collection by
way of a specialized role and mapping to a collection's storage layer.
External users (e.g., analysts) can discover
and then obtain access to safe collections by pulling lists of available safe collections and requesting
access to individual collections.
Stewards for a safe collection are presented with a list of user requests and must explicitly approve access
for a specific duration. Throughout the duration of a user's access to a collection
stewards may review their actions and revoke access at any time. When users attempt
to export data from an export controlled safe-collection, the data exported data is not immediately
accessible to the user. Rather, the exported data is collected and stored privately under the assumption that it contains microdata, and is only released after being reviewed by a steward.

@@

Apollo, an analyst, authenticates to \NAME and requests access to a private safe collection.
The request is added to the stewarship dashboard and is visible to all stewards
of that collection. Daphnie, a steward for the requested
safe collection views and then approves Apollo's request for a period of $n$ days.
The approval process adds the appropriate policies to the
safe collection granting access to Apollo. Apollo can now submit analysis tasks through the various
interfaces offered by \NAMENS, that specify data items from the safe collection as inputs.
The REST API marks the analysis task as export controlled since it draws results
from an export controlled safe collection.
The submitted task is posted to an internal queue from which a
worker node in the \NAME private subnet picks up Apollo's task. The worker node assumes Apollo's identity and, with
his credentials, requests data from the safe collection. Since Apollo now has read privileges for the
safe collection, the data is fetched to the worker node on which the analysis task is executed.
Upon completion, the worker node puts the generated results in the export-controlled
output directory of the safe collection since these results are generated from export controlled data.
Apollo's analysis is now complete, but he is unable to view the
results as they require steward approval. This flow is show in \figurename~\ref{fig:flow1}

@@

Once Apollo's analysis task is complete, Apollo realizes that he requires export approval before he may view the
results, so he issues a request to see his results. Daphnie, one of the collection's stewards, receives
the export request on her dashboard. Daphnie assumes
the steward role with a fresh authenticated token and accesses the results and the complete description
of Apollo's analysis task. Daphnie reviews the data and the analysis to determine if the results are
safe for export. When she is satisfied with Apollo's process, she initiates data export. \NAME
uses Daphnie's temporary credentials to transfer the results from the safe collection to a public
safe collection, accessible only to Apollo, and updates the task information with the new location of the results. Apollo can now log on
and view or download his results. This flow is show in \figurename~\ref{fig:flow2}

@@

Data derived from safe collections (e.g., results from tasks that consumed data from that collection)
are stored in the same safe collection, thus inheriting the same policies as the safe collection itself.
As a result, only stewards can view the derived data and determine if it is safe to
export the data
Necessarily, stewards have complete visibility over all data that was used as input to an analysis, as
well as the operations (analysis steps) made by the analysts to derive their results.
It is, therefore, possible to trace the entire provenance of derived data from an
analysis pipeline given that all analysis tasks are logged.
If an export request is approved, data is copied from the host collection to a public safe collection.
It is important to note that even public safe collections have server side encryption and
are only accessible to authenticated users via temporary signed URLs.
As derived data are stored within the parent's safe collection, users can continue
to perform analysis operations on results without oversight,
and only request export of data that must leave the safe collection (e.g., the
final results of a deep analysis pipeline).

>>>>

Selecting the policies, protocols, infrastructure, and
software settings to ensure that individual datasets are adequately secured requires consideration
on a case-by-case basis.
%To represent these policies, such that dataset owners can manage these considerations,
%we define safe-collections.
Safe collections enable system administrators to partition the key administrative
aspects of managing the relationship between data and users (e.g., analysts who use those data).
Safe collections encapsulate the data-store in which a dataset resides, and a consistent set of policies
that controls access to that dataset.

@@

\textbf{SecurityClass:} A binary classification that specifies the constraints
on accessing data in that collection. Setting the SecurityClass to \texttt{Public} means that the data are freely
(but securely) accessible to a set of permitted users and no other
data administration policies can be set.  Setting SecurityClass to \texttt{Private}
requires that safe collection policies are then set.

\textbf{Classification:} Categorizes the type of data stored in the
safe collection, including specifying that the data are controlled by
a license or that the data are subject to regulations (e.g., HIPAA or FEDRAMP).
This information is referenced and used by stewards and administrators when
managing a safe collection.

@@

Stewards are responsible for activities such as approving access to a collection, reviewing
export requests, and managing policies on the collection.
As described above, stewards are pre-associated with a particular safe collection by
way of a specialized role and mapping to a collection's storage layer.
External users (e.g., analysts) can discover
and then obtain access to safe collections by pulling lists of available safe collections and requesting
access to individual collections.
Stewards for a safe collection are presented with a list of user requests and must explicitly approve access
for a specific duration. Throughout the duration of a user's access to a collection
stewards may review their actions and revoke access at any time. When users attempt
to export data from an export controlled safe-collection, the exported data are not immediately
accessible to the user. Rather, the exported data are collected and stored privately under the assumption that they contain microdata, and are only released after being reviewed by a steward.

@@

Apollo, an analyst, authenticates to \NAME and requests access to a private safe collection.
The request is added to the stewarship dashboard and is visible to all stewards
of that collection. Daphnie, a steward for the requested
safe collection views and then approves Apollo's request for a period of $n$ days.
The approval process adds the appropriate policies to the
safe collection granting access to Apollo. Apollo can now submit analysis tasks through the various
interfaces offered by \NAMENS, that specify data items from the safe collection as inputs.
The REST API marks the analysis task as export controlled since it draws results
from an export controlled safe collection.
The submitted task is posted to an internal queue from which a
worker node in the \NAME private subnet picks up Apollo's task. The worker node assumes Apollo's identity and, with
his credentials, requests data from the safe collection. Since Apollo now has read privileges for the
safe collection, the data are fetched to the worker node on which the analysis task is executed.
Upon completion, the worker node puts the generated results in the export-controlled
output directory of the safe collection since these results are generated from export controlled data.
Apollo's analysis is now complete, but he is unable to view the
results as they require steward approval. This flow is show in \figurename~\ref{fig:flow1}.

@@

Once Apollo's analysis task is complete, Apollo realizes that he requires export approval before he may view the
results, so he issues a request to see his results. Daphnie, one of the collection's stewards, receives
the export request on her dashboard. Daphnie assumes
the steward role with a fresh authenticated token and accesses the results and the complete description
of Apollo's analysis task. Daphnie reviews the data and the analysis to determine if the results are
safe for export. When she is satisfied with Apollo's process, she initiates data export. \NAME
uses Daphnie's temporary credentials to transfer the results from the safe collection to a public
safe collection, accessible only to Apollo, and updates the task information with the new location of the results. Apollo can now log on
and view or download his results. This flow is show in \figurename~\ref{fig:flow2}.

@@

Data derived from safe collections (e.g., results from tasks that consumed data from that collection)
are stored in the same safe collection, thus inheriting the same policies as the safe collection itself.
As a result, only stewards can view the derived data and determine if it is safe to
export the data
Necessarily, stewards have complete visibility over all data that was used as input to an analysis, as
well as the operations (analysis steps) made by the analysts to derive their results.
It is, therefore, possible to trace the entire provenance of derived data from an
analysis pipeline given that all analysis tasks are logged.
If an export request is approved, data are copied from the host collection to a public safe collection.
It is important to note that even public safe collections have server side encryption and
are only accessible to authenticated users via temporary signed URLs.
As derived data are stored within the parent's safe collection, users can continue
to perform analysis operations on results without oversight,
and only request export of data that must leave the safe collection (e.g., the
final results of a deep analysis pipeline).

### 2017-08-15 introduction.tex
data are plural

At the same time, data sensitivity and security concerns are well founded. While many datasets are public, the most
valuable ones are often considered `private' or `protected' due to legal or contractual obligations under which the data
is shared, or even, the most extreme cases, under governmental regulation. However, the limited and proprietary nature
of particular datasets can further increase perceived value to researchers given that the space of possible
questions and possible answers that can be derived from the data is, in a sense, hidden from competition.
Simultaneously, increasing sensitivity adds to the burden of ensuring data security.

>>>>

At the same time, data sensitivity and security concerns are well founded. While many datasets are public, the most
valuable ones are often considered `private' or `protected' due to legal or contractual obligations under which the data
is shared, or even, the most extreme cases, under governmental regulation. However, the limited and proprietary nature
of particular datasets can further increase perceived value to researchers given that the space of possible
questions and possible answers that can be derived from the data are, in a sense, hidden from competition.
Simultaneously, increasing sensitivity adds to the burden of ensuring data security.

### 2017-08-15 relatedwork.tex
added ref to foster paper and fixed various minor typos
data are plural

NORC~\cite{lane2008using}, a research institution at the University of Chicago, operates a
data enclave to support researchers' investigations into programmatic, business, and policy decisions. 
NORC stores and manages a broad set of sensitive data in their FISMA compliant data enclave. The enclave
was designed with the vision of supporting multi-researcher collaboration via remote access terminals.
All microdata is hosted in a secure NORC server. 
This common approach to developing enclaves ensures security by access control, and limiting computation
to the server. However, it is limited by storage and computational infrastructure and the need to
host and operate local infrastructure. Moreover, it is typically not well
suited for large scale data analysis. Furthermore, users could potentially read microdata through 
their terminal. Thus, this approach only guarantees the privacy of the bulk of microdata.

@@

The National Center for Health Statistics (NCHS) Research Data Center (RDC) \cite{cdc} hosts a large collection of
restricted datasets. These datasets contain health information and are subject to HIPAA guidelines.
The data is accessible on-premise at the NCHS RDC or the Federal Statistical RDC, or in some
cases via remote access. 
%several datasets are not available for remote access. 
RDC restricts access to direct identifiers such as name and 
social security number while leaving indirect identifiers such as geography accessible. Even with the constraints
on access and limited record accessibility, access to potentially identifying information leaves this system open to
linkage attacks. This is mitigated, to some extent, by strong vetting of research proposals.

@@

>>>>

NORC~\cite{lane2008using}, a research institution at the University of Chicago, operates a
data enclave to support researchers' investigations into programmatic, business, and policy decisions. 
NORC stores and manages a broad set of sensitive data in their FISMA compliant data enclave. The enclave
was designed with the vision of supporting multi-researcher collaboration via remote access terminals.
All microdata are hosted in a secure NORC server. 
This common approach to developing enclaves ensures security by access control, and limiting computation
to the server. However, it is limited by storage and computational infrastructure and the need to
host and operate local infrastructure. Moreover, it is typically not well
suited for large scale data analysis. Furthermore, users could potentially read microdata through 
their terminal. Thus, this approach only guarantees the privacy of the bulk of microdata.

@@

The National Center for Health Statistics (NCHS) Research Data Center (RDC) \cite{cdc} hosts a large collection of
restricted datasets. These datasets contain health information and are subject to HIPAA guidelines.
The data are accessible on-premise at the NCHS RDC or the Federal Statistical RDC, or in some
cases via remote access. 
%several datasets are not available for remote access. 
RDC restricts access to direct identifiers such as name and 
social security number while leaving indirect identifiers such as geography accessible. Even with the constraints
on access and limited record accessibility, access to potentially identifying information leaves this system open to
linkage attacks. This is mitigated, to some extent, by strong vetting of research proposals.

@@

Foster~\cite{foster-stakeholder} reviews approaches to enabling secure analysis of sensitive data,
and proposes---but does not describe an implementation of---a cloud architecture with similarities to that described here.

### 2017-08-15 summary.tex
Section for summary
data are plural

There are significant opportunities for new scientific discoveries to be made
through the collaborative analysis of sensitive and protected
data. While many approaches to storing such data
have been adopted in production deployments, in most cases there are
considerable barriers that overly restrict data access and analysis. 
Importantly, there is little functionality that enales dataset owners and administrators
to define and apply sophisticated administrative models and policies 
for securely managing that data.

To address these concerns we have developed two important abstractions
that allow data owners to securely manage data in ways that permit
controlled access and analysis. The `secure collection' provides a 
policy-based approach for defining possible interactions with a particular
dataset. `Data stewards' are associated with collections to then
manage the collections and review the actions of their users. These
constructs have been implemented in the \NAME data enclave
and have since been adopted to protect a range of sensitive
social science data. Initial experiences with the model
have been positive, with data owners valuing the ability 
to securely manage collections, and users valuing the clearly
defined policies for data access and analysis.

>>>>

\section{Summary}

There are significant opportunities for new scientific discoveries to be made
through the collaborative analysis of sensitive and protected
data. While many approaches to storing such data
have been adopted in production deployments, in most cases there are
considerable barriers that overly restrict data access and analysis. 
Importantly, there is little functionality that enables dataset owners and administrators
to define and apply sophisticated administrative models and policies 
for securely managing those data.

To address these concerns we have developed two important abstractions
that allow data owners to securely manage data in ways that permit
controlled access and analysis. The `safe collection' provides a 
policy-based approach for defining possible interactions with a particular
dataset. `Data stewards' are associated with collections to then
manage the collections and review the actions of their users. These
constructs have been implemented in the \NAME data enclave
and have since been adopted to protect a range of sensitive
social science data. Initial experiences with the model
have been positive, with data owners valuing the ability 
to securely manage collections, and users valuing the clearly
defined policies for data access and analysis.
